---
title: "Cohort models 1"
subtitle: "EPIB  676 session 4, McGill University"
author: "Alton Russell"
date: "13 Jan 2023"
format: revealjs
editor: visual
---

## Today

-   **Discrete-time cohort state-transition models**

-   Matrix solution

-   Cohort simulation

## Packages

```{r}
#| echo: true
library(heemod)

```

## **Cohort state-transition models**

-   Dynamic mathematical model

-   A hypothetical cohort of individuals transition between health states over time.

-   Classical example is the Markov model, where transition probabilities

    -   Are constant

    -   Depend only on your current state

## When to use?

-   Decision problem has a dynamic component

    -   Disease process can vary over time

-   Can be described with a reasonable number of health states

## **States**

-   Mutually exclusive and exhaustive

-   Individuals in the same state treated as identical

    -   same costs, same health outcomes

    -   same probability of transitioning out (no matter how long they've been there)

-   **Possible state names:** healthy, acute infection, chronic infection, cirrhosis, untreated disease, in treatment, on wait list, lost to follow-up, dead...

## Transitions, cycles, time horizon

-   Only transition at end of a cycle

-   Cycle duration is a design choice (second, day, month, quarter, year...)

-   Time horizon = cycle duration x number of cycles

## Transition matrix

::: columns
::: {.column width="50%"}
$$
\begin{bmatrix}
p_{s_1 \rightarrow s_1} & p_{s_1 \rightarrow s_2} & p_{s_1 \rightarrow s_3}\\
p_{s_2 \rightarrow s_1} & p_{s_2 \rightarrow s_2} & p_{s_2 \rightarrow s_3}\\
p_{s_3 \rightarrow s_1} & p_{s_3 \rightarrow s_2} & p_{s_3 \rightarrow s_3}
\end{bmatrix}
$$
:::

::: {.column width="50%"}
-   $n_s \times n_s$ matrix of transition probabilities

-   Rows correspond to "**from**" state (sum to 1)

-   Columns correspond to "**to**" states
:::
:::

## Transition matrix example

::: columns
::: {.column width="40%"}
```{r}
#| echo: true
#| eval: false
mat_dim <- 
  define_transition(
    state_names = 
      c('A', 'B', 'C'),
   .3, .6, .1, 
   .2, .5, .3,
    0,  0,  1)

plot(mat_dim)
```

::: {.column width="50%"}
$$
\begin{array}{c}\begin{matrix}a & b & c \end{matrix} 
\\\left[\ 
\begin{matrix}
0.3 & 0.6 & 0.1\\
0.2 & 0.5 & 0.3\\
0 & 0 & 1
\end{matrix}\ \right]
\end{array}
$$
:::
:::

::: {.column width="60%"}
```{r}
#| fig-width: 6
#| fig-height: 6
mat_dim <- define_transition(
  state_names = 
  c('A', 'B', 'C'),
   .3, .6, .1, 
   .2, .5, .3,
    0,  0,  1)
plot(mat_dim)

```
:::
:::

## Absorbing states

::: columns
::: {.column width="50%"}
```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 3
mat_dim <- 
  define_transition(
    state_names = 
      c('Alive', 'Dead'),
   .9, .1,
   0, 1)
plot(mat_dim)
```

$$
\begin{array}{c}\begin{matrix}alive & dead \end{matrix} 
\\\left[\ 
\begin{matrix}
0.9 & 0\\
0 & 1
\end{matrix}\ \right]
\end{array}
$$
:::

::: {.column width="50%"}
-   No escape

-   Death is common example

-   Transition matrix row has a 1 on diagonal entry, all other entries 0
:::
:::

## Tunnel states

::: columns
::: {.column width="50%"}
```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 3
mat_dim <- 
  define_transition(
    state_names = 
      c('HIV-', 'Infected', 'HIV+'),
   .9, .1, 0,
   0, 0, 1,
   0, 0, 1)
plot(mat_dim)
```

$$
\begin{array}{c}\begin{matrix}HIV- & Infected & HIV+\end{matrix} 
\\\left[\ 
\begin{matrix}
0.9 & .1 & 0\\
0 & 0 & 1\\
0 & 0 & 1
\end{matrix}\ \right]
\end{array}
$$
:::

::: {.column width="50%"}
-   Temporary state, 1 cycle only

-   A 'workaround' for the Markov property.

-    If costs, outcomes, or transitions differ for beginning or end of health condition, break into tunnel states
:::
:::

## Tunnel state example

## Initial state vector $m_0$

-   $1 \times n_s$ vector showing how cohort is distributed at time 0

-   Sum of the vector is the total cohort size

-   Everyone can be in one state

    $$
    m_0 =
    \begin{bmatrix}
    10,000 & 0 & 0 & 0\\
    \end{bmatrix}
    $$

-   Or spread across the states

$$
m_0 = 
\begin{bmatrix}
2,000 & 3,000 & 1,000 & 4,000\\
\end{bmatrix}
$$

## Rates and probabilities

-   **Rates** are instintaneous force of an event

-   **Probabilities** is the cumulative risk over a defined period

-   We need: transition probabilities specific to cycle length

-   We may have:

    -   probability for different cycle length (2-year instead of 6 month)

    -   rate (e.g., 200 events per person-year)

## Converting rates to probabilities

Generating an annual probablity $p_{year}$ from an event rate $r$ with units events/year:

$$
p_{year} = 1 - exp(-r) \quad r = log(
$$

Example:

-   In a population of 800, 87 became ill over a 20 year period

-   $r = 87/(2 \times 800) = 0.0544$ illnesses/year

-   $p_{year} = 1 - exp(-0.0544) = 0.0529$

## Prob → rate → prob

-   If probability is for a duration that doesn't match your cycle length, you **can not** simply multiply/divide

-   Instead convert to a rae

## Time horizon

## Closed and open populations

## 

## Today

-   Discrete-time cohort state-transition models

-   **Matrix solution**

-   Cohort simulation

## Today

-   Discrete-time cohort state-transition models

-   Matrix solution

-   **Cohort simulation**

## Recap

## Logistics

-   Assignment 1 (cost-effectiveness, decision trees) due Wednesday 1/18
