---
title: "Assignment 1: Decision trees, economic evaluation"
author: "YOUR NAME HERE"
format: html
editor: visual
---

## Part 1: Decision trees

In this problem set, you will develop a decision analysis to help determine whether a newer clip should be used to close bleeding peptic ulcers in the gastrointestinal tract. The newer clips are called over-the-scope clips, abbreviated OTSc. Randomized trial data show that rebleeding rates are lower following endoscopic closure with OTSc as compared to when standard therapy clips are used, but OTSc are also significantly more expensive. You will use decision analysis to determine whether OTSc are 'worth' the added expense, either as a first line therapy (i.e. to treat all peptic ulcer bleeds initially) or only for rebleeds (i.e. only if a standard therapy clip was used and failed).

We assume all peptic ulcer bleeds are fully resolved during a short hospitalization and therefore our analysis uses a time horizon of only 30 days. No discounting is needed due to the short time horizon. Because this analysis involves very small differences in utility, quality-adjusted life days (QALDs) are used instead of quality-adjusted life years (QALYs).

```{r}


```

## Part 2: Economic evaluation

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
intervention_A <- list(
  intervention_cost = 20, #dollars
  downstream_cost = 300, #dollars
  downstream_effect = 21 #Quality-adjusted life years
)
```
